---
title: "Financial Analysis"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "Adel Ben Snoussi, El-Amine Maamar, Eric Brandstätter, Nathanaël G. Dürst"
output: html_document
---

```{r, echo=F, results="hide", message=F, warning=F}

#Load library
library("quantmod")
library("PerformanceAnalytics")

#List of the companies analyzed
tickerList = c("AAPL", "CSCO", "INTC", "HD", "GOOG", "JPM")

#Import data
getSymbols.yahoo(
  tickerList,
  env = globalenv(),
  from = "2020-01-01",
  to = "2021-01-01",
  periodicity = "daily"
)

```

## Introduction

In this financial analysis, we will discuss about the stocks of **Apple**, **Cisco Systems**, **Google**, **JPMorgan Chase**, **Home Depot** and **Intel**.

## Some graphs

### Apple

The closing price of Apple stocks has increased a lot over the last year.

```{r, echo=F, results="hide", message=F, warning=F}

#Plot AAPL's closing price
plot(
  AAPL$AAPL.Close,
  main="AAPL Closing Price over the year",
  ylab="Closing price in USD",
)

```

### Cisco Systems

The closing price of Cisco stocks has decreased over the past year.


```{r, echo=F, results="hide", message=F, warning=F}

#Plot CSCO's closing price
plot(
  CSCO$CSCO.Close,
  main="CSCO Closing Price over the year",
  ylab="Closing price in USD",
  col = "blue",
)

```

It moved a lot but did not progress in the end.

### Google

Here you can see a candle chart of Google's stock over the last quarter of 2020.

```{r, echo=F, results="hide", message=F, warning=F}

#Candle chart GOOG 2020 Q4
candleChart(
  GOOG,
  subset = "2020-10/2020-12",
  name="GOOG Candle chart 2020 Q4",
)

```

### JPMorgan Chase

We can clearly see a volume spike on JPM's candle chart in november of 2020.

```{r, echo=F, results="hide", message=F, warning=F}

#Candle chart JPM 2020 Q4
candleChart(
  JPM,
  subset = "2020-10/2020-12",
  name="JPM Candle chart 2020 Q4",
)

```

```{r, echo=F, results="hide", message=F, warning=F}

#--------------- [START] Mean and Variance Loop ---------------#

########## This needs to be modified to use a matrix to store data (cleaner code on the recommendation of Lionel)

for (tickerName in tickerList) {

  #Variables' name generation
  varNameRe <- paste("returns",tickerName, sep = "")
  varNameMe <- paste("mean",tickerName, sep = "")
  varNameVa <- paste("var",tickerName, sep = "")

  #Computations
  returnsStock <- na.omit(ClCl(get(tickerName)))
  names(returnsStock) <- tickerName
  meanStock <- mean(returnsStock)
  varStock <- as.numeric(var(returnsStock))
  
  #Assigning values to their variables
  assign(varNameRe, returnsStock)
  assign(varNameMe, meanStock)
  assign(varNameVa, varStock)

}

#Cleaning environment after loop end
rm(returnsStock, varStock, meanStock, tickerName, varNameMe, varNameRe, varNameVa)

#--------------- [END] Mean and Variance Loop ---------------#

```

## Means and variances

Here we will talk about the mean and variance of all of the stocks' closing prices we are analyzing for the year 2020.

Google's mean is **`r meanGOOG`** and its variance is **`r varGOOG`**.

On the other hand, Apple's mean is **`r meanAAPL`** and its variance is **`r varAAPL`**.

Home Depot have different values: mean = **`r meanHD`** & variance = **`r varHD`**.

Whereas Intel has a mean of **`r meanINTC`** and a variance of **`r varINTC`**.

JPM has a mean of **`r meanJPM`** and a variance of **`r varJPM`**.

Last but not least, Cisco's mean is equal to **`r meanCSCO`** with a variance of **`r varCSCO`**

```{r, echo=F, message=F, warning=F}

mean_var_mat <- matrix(NA, 2, 6)
dimnames(mean_var_mat)[[1]] <- c("Expected value", "Variance")
dimnames(mean_var_mat)[[2]] <- c("AAPL", "CSCO", "INTC", "HD", "GOOG", "JPM")
mean_var_mat[1, ] <- c(meanAAPL, meanCSCO, meanINTC, meanHD, meanGOOG, meanJPM)
mean_var_mat[2, ] <- c(varAAPL, varCSCO, varINTC, varHD, varGOOG, varJPM)
knitr::kable(mean_var_mat)
```

```{r, echo=F, results="hide", message=F, warning=F}

#--------------- [START] Sharpe ratio ---------------#

#Import data

getSymbols.yahoo(
  tickerList,
  env = globalenv(),
  from = "2011-01-01",
  to = "2021-01-01",
  periodicity = "daily"
)

#--------------- [START] Sharpe ratio Loop ---------------#

########## This needs to be modified to use a matrix to store data (cleaner code on the recommendation of Lionel)

riskFree <- 0.03/252

for (tickerName in tickerList) {

  #Variables' name generation
  varNameS <- paste("sharpe",tickerName, sep = "")

  #Computations
  dailyReturnStock <- dailyReturn(get(tickerName))
  sdStock <- sd(dailyReturnStock)
  excessReturnStock <- dailyReturnStock - riskFree
  sharpeStock <- mean(excessReturnStock)/sdStock*sqrt(252)

  #Assigning values to their variables
  assign(varNameS, sharpeStock)

}

#Cleaning environment after loop end
rm(varNameS, dailyReturnStock, sdStock, excessReturnStock, sharpeStock, tickerName)

#--------------- [END] Sharpe ratio Loop ---------------#

#--------------- [END] Sharpe ratio ---------------#

```

## Sharpe ratio

The lowest and highest Sharpe ratio are:`r summary(c(sharpeAAPL, sharpeGOOG, sharpeHD, sharpeJPM, sharpeINTC, sharpeCSCO))[-2:-5]`

They belong to **Cisco and Apple** respectively.

Between them, we find the others : `r sort(c(sharpeGOOG, sharpeHD, sharpeJPM, sharpeINTC))` (INTC, JPM, GOOG, HD).

```{r, echo=F, message=F, warning=F}
yearly_sharpe = cbind(sharpeAAPL, sharpeCSCO, sharpeINTC, sharpeHD,sharpeGOOG, sharpeJPM)
colnames(yearly_sharpe) = c("AAPL", "CSCO", "INTC", "HD", "GOOG", "JPM")

knitr::kable(yearly_sharpe)
```

```{r, echo=F, results="hide", message=F, warning=F}

#--------------- [START] Weighted portfolio - Uses the first set of data for comparison ---------------#

weightPortfolio <- rep((1/length(tickerList)), times = length(tickerList))

returnsAll <- NULL

########## This needs to be modified to use a matrix to store data (cleaner code on the recommendation of Lionel)

for (tickerName in tickerList) {

  dailyReturnStock <- dailyReturn(get(tickerName), subset = "2020-01-01/2021-01-01")
  returnsAll <- merge.xts(returnsAll, dailyReturnStock)
}

names(returnsAll) <- tickerList

#Cleaning environment after loop end
rm(dailyReturnStock, tickerName)

returnsPortfolio <- Return.portfolio(returnsAll, weights = weightPortfolio)
names(returnsPortfolio) <- "Portfolio"

varPortfolio <- as.numeric(var(returnsPortfolio))

#--------------- [END] Weighted portfolio ---------------#

```

## Equally weighted portfolio

```{r, echo=F, results="hide", message=F, warning=F}

#--------------- [START] Compute optimum portfolio ---------------#

AAPL_return = dailyReturn(AAPL)
CSCO_return = dailyReturn(CSCO)
INTC_return = dailyReturn(INTC)
HD_return = dailyReturn(HD)
GOOG_return = dailyReturn(GOOG)
JPM_return = dailyReturn(JPM)

mat_return = cbind(AAPL_return,CSCO_return, INTC_return, HD_return, GOOG_return, JPM_return)
colnames(mat_return)= c("AAPL", "CSCO", "INTC", "HD", "GOOG", "JPM")

mu = apply(mat_return, 2, mean)
Sigma <- var(mat_return)

# equally weighted portfolio
equal_return = (AAPL_return + CSCO_return + INTC_return + HD_return + GOOG_return + JPM_return)/6

mu_equal = mean(equal_return)
var_equal = var(equal_return)
```

The individual variance of these 6 stocks are:

```{r, echo=F}
diag_Sigma_mat = cbind(diag(Sigma)[[1]][1],diag(Sigma)[[2]][1],diag(Sigma)[[3]][1],diag(Sigma)[[4]][1],diag(Sigma)[[5]][1],diag(Sigma)[[6]][1])
colnames(diag_Sigma_mat) = c("AAPL", "CSCO", "INTC", "HD", "GOOG", "JPM")
knitr::kable(diag_Sigma_mat)
```

And the variance of the equally weighted portfolio is `r var_equal` which is smaller than the variance of the indivual return of any stocks.

## Optimum portfolio

The optimally weighted portfolio (made of these 6 stocks) is called "Min_var".

The equally weighted portfolio (made of these 6 stocks) is called "Equal".

```{r, echo=F, message=F, warning=F}

var = rep(1,6)
inv_Sigma = solve(Sigma)
P = as.numeric( inv_Sigma %*% var / as.numeric(t(var) %*% inv_Sigma %*% var))

optimum_return = AAPL_return*P[1] + CSCO_return*P[2] + INTC_return*P[3] + HD_return*P[4] + GOOG_return*P[5] + JPM_return*P[6]
min_var = apply(optimum_return, 2, var)
mean_optimum_return = apply(optimum_return, 2, mean)

investment_summary <- matrix(NA, 2, 8)
dimnames(investment_summary)[[1]] <- c("Expected value", "Variance")
dimnames(investment_summary)[[2]] <- c("AAPL", "CSCO", "INTC", "HD", "GOOG", "JPM", "Min_var", "Equal")
investment_summary[1, ] <- c(mu, mean_optimum_return, mu_equal)
investment_summary[2, ] <- c(diag(Sigma), min_var, var_equal)
knitr::kable(investment_summary)

plot((sqrt(investment_summary[2, ])), investment_summary[1, ],
     main="Stocks Portfolio",
     xlab="Daily Investment Standard Deviation",
     ylab="Daily Expected Value of Investment ",
     xlim = c(min(sqrt(investment_summary[2, ])), max(sqrt(investment_summary[2, ])+1e-04)),
     ylim = c(0, 0.0012),
     lty = "solid",
     lwd = 2, 
     col = 1:8,
     cex = 1,
     pch = 19,
grid())
text(sqrt(investment_summary[2, ]), investment_summary[1, ], labels = names(investment_summary[2, ]), cex = 0.5, pos = 3)

#--------------- [END] Compute optimum portfolio ---------------#

```

As we may notice from the graph, the equally weighted portfolio performs well in a way that it offers both upside potential with relatively high investment output and downside protection with low investment risk.
But the optimum portfolio performs better with lower investment risk and higher returns.

In this latter case, the respective weights of each stocks "AAPL", "CSCO", "INTC", "HD", "GOOG", "JPM" are : 
```{r, echo=F, message=F, warning=F}
weights = cbind(P[1],P[2],P[3],P[4],P[5],P[6])
colnames(weights) = c("AAPL", "CSCO", "INTC", "HD", "GOOG", "JPM")

knitr::kable(weights)
```